Effect of multiple reflection of the sky signal off the HERA element on the measurements of 21cm power spectrum

### 2015-11-06 DishReflectometry/Paper1_v0.tex
adding the file names of reflectometry data.

\documentclass[12pt,preprint]{aastex}
\usepackage{float,amsmath}
%\usepackage[margin=1in]{geometry}
%\usepackage{titlesec} %used to format titles
\usepackage{graphicx} %for handling figures
%\usepackage[none]{hyphenat} %disallows hyphenated words
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\volt}{{v}}
\newcommand{\vis}{{V}}
\newcommand{\sky}{{\rm sky}}
\newcommand{\bmvolt}{{a}}
\newcommand{\bm}{{A}}
\newcommand{\thhat}{{\hat\theta}}
\newcommand{\fngexp}{{e^{\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\ifngexp}{{e^{-\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\myemail}{skywalker@galaxy.far.far.away}

@@

Consider a two element interferometer with a baseline $\vec b$ and antenna field pattern $A(\thhat, \nu)$. If we denote the electric field sky signal in direction $\thhat$ by $\volt_\sky(\vec \theta, \nu)$ and the receiver noise of each element as $\volt_{rec}(\nu)$ then the voltage output of each antenna may be written as,  
\begin{equation}
\volt_{1}(\thhat, \nu) = \bmvolt_{1}(\thhat,\nu)\volt_\sky(\thhat,\nu)\nonumber\\
\end{equation}
\begin{equation}
\volt_{2} (\thhat, \nu)= \bmvolt_{2}(\thhat,\nu)\volt_\sky(\thhat,\nu)\fngexp\nonumber\\
\end{equation}
where $\bmvolt_{1}(\thhat,\nu)$ and $\bmvolt_{2}(\thhat,\nu)$ are the electric field response patterns 
of the two antennas.
Hence, the time averaged visibility measured by the interferometer may be written as, 
\begin{equation}
\vis(\vec b, \nu) =  \int  \volt_{1}(\thhat,\nu)  \volt_{2}^{*} (\thhat, \nu) \ifngexp d\Omega 
\label{eq1}
\end{equation}

@@

Consider again a two element interferometer as described in section 2. Two antennas are assumed to be exactly identical so that their electrical parameters are same. We denote the voltage reflection coefficient of each feed by $\Gamma_{f}$. The voltage reflection coefficient of the reflector dish is denoted by $\Gamma_{d}$. Upon first incidence of the sky signal onto the dish, it is reflected from the entire dish and is focussed onto the feed. Hence, for first incidence $\Gamma_{d}=1$. Sky signal that would return to the dish after successive reflections off the feed and the dish is only reflected from a small area around the vertex of the dish and therefore, $\Gamma_{d}<1$. In our analysis, we consider one reflection of the sky power from the feed and subsequent reflection off the dish as one single reflection. If the focal length of the dish is $l$, assuming the feed is located at the focus, the roundtrip delay after first reflection will be $\Delta \tau = {2l \over c}$.

\subsection{Effect of first order reflections on visibility and delay spectrum}
Upon first incidence of the sky signal $V_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
V_{1} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  (1-\Gamma_{f})^{2}\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber\\
\label{eq4}
\end{eqnarray}
\begin{eqnarray}
V_{2} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} (1-\Gamma_{f})^{2}exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  \right]exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
\label{eq5}
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. 
The two antennas are assumed to have identical beam pattern which is $A_{1}=A_{2} =A$. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna response $A$ and denote them as $A_{f}= A(1-\Gamma_{f})$ here after. Hence, the complex visibility could be,

\begin{eqnarray}
V e^{i\phi}(\nu) & = & \int V_{1}V_{2}^{*} d\Omega \nonumber\\
		        & = & \int |A_{f}|^{2}|V_{sky}|^{2} |\left [ 1+(1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  \right]|^{2} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega
\label{eq6}		        
\end{eqnarray}

 Furthermore, combining all the terms that contain instrument response we define a common instrument parameter, written as, 
\begin{eqnarray}
\Sigma(\vec \theta, \nu) & = & |A_{f}|^{2} |\left [ 1+ (1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]|^{2}\nonumber\\
 	    & = &  |A_{f}|^{2}  \left[ 1+ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} + 2 \Re \left ( (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} e^{2\pi i\nu \Delta \tau} \right) \right] 
\label{eq{7}}	    
\end{eqnarray}

@@

Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
 Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay is centred at a delay $\tau = \tau_{g}$ in the delay spectrum whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\tau_{g} +\Delta \tau$. Therefore, the correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal thus extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected. Also, any reflections arriving at the feed at any non-zero  value of the delay $\Delta \tau$ which could be the result of reflections from the feed structures would also generate non zero delay response at those delays. 
If the dish vertex is so constructed such that the reflection from the dish vertex is negligible i.e, $\Gamma_{d}<<1\ or = 0$ then the second and the third term of the $\Sigma$ vanish. Also, if we assume a $\Gamma_{a} =-10dB$ and $\Gamma_{d}= -20dB$, the second term in the expression of $\Sigma(\vec \theta, \nu)$ becomes negligible compared to the first and the third term of the expression and, therefore, can be neglected. \\

\subsection{Effect of higher order reflection on the delay spectrum} 
Correlation between the higher order reflections of the signal in any antenna with the direct component of the other would result in a delay response at integral multiple of $\Delta \tau$ and vice versa. 
If $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{equation}
V_{1} =  A V_{sky}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}
and,

\begin{equation}
V_{2} = A V_{sky}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

where $\Phi = 2\pi\nu\Delta \tau$. The relevant terms in the corresponding visibility and subsequently in the delay spectrum would be the ones which represent the correlation between the direct component of one antenna with the reflected components of the other resulting in higher order terms  in $\Sigma(\vec \theta, \nu)$. The corresponding instrument response may be written as,

\begin{equation}
\Sigma(\vec \theta, \nu) = A_{f} \left [ 1+ 2 \Re \left [(1-\Gamma_{f})( \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}) \right ]\right ]
\end{equation}

and the measured visibility spectrum, integrated over all sky would be, 
\begin{equation}\label{eqn:series1}
V_{meas}  = I_{sky}A_{f}\Sigma( \nu)
\end{equation}

>>>>

\documentclass[12pt,preprint]{aastex}
\usepackage{float,amsmath}
%\usepackage[margin=1in]{geometry}
%\usepackage{titlesec} %used to format titles
\usepackage{graphicx} %for handling figures
%\usepackage[none]{hyphenat} %disallows hyphenated words
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\volt}{{v}}
\newcommand{\vis}{{V}}
\newcommand{\sky}{{\rm sky}}
\newcommand{\bmvolt}{{a}}
\newcommand{\bm}{{A}}
\newcommand{\thhat}{{\hat\theta}}
\newcommand{\fngexp}{{e^{\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\ifngexp}{{e^{-\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\dfngexp}{{e^{2\pi i\nu \Delta \tau}}}
\newcommand{\myemail}{skywalker@galaxy.far.far.away}

@@

Consider a two element interferometer with a baseline $\vec b$ and antenna field pattern $A(\thhat, \nu)$. If we denote the electric field sky signal in direction $\thhat$ by $\volt_\sky(\vec \theta, \nu)$ and the receiver noise of each element as $\volt_{rec}(\nu)$ then the voltage output of each antenna may be written as,  
\begin{equation}
\volt_{1}(\thhat, \nu) = \bmvolt_{1}(\thhat,\nu)\volt_\sky(\thhat,\nu)\nonumber\\
\end{equation}
\begin{equation}
\volt_{2} (\thhat, \nu)= \bmvolt_{2}(\thhat,\nu)\volt_\sky(\thhat,\nu)\dfngexp\nonumber\\
\end{equation}
where $\bmvolt_{1}(\thhat,\nu)$ and $\bmvolt_{2}(\thhat,\nu)$ are the electric field response patterns 
of the two antennas.
Hence, the time averaged visibility measured by the interferometer may be written as, 
\begin{equation}
\vis(\vec b, \nu) =  \int  \volt_{1}(\thhat,\nu)  \volt_{2}^{*} (\thhat, \nu) \ifngexp d\Omega 
\label{eq1}
\end{equation}

@@

Consider again a two element interferometer as described in section 2. Two antennas are assumed to be exactly identical so that their electrical parameters and the field pattern are identical i.e, $a_{1}(\thhat,\nu) = a_{2}(\thhat,\nu)$. We denote the voltage reflection coefficient of each feed by $\Gamma_{f}$. The voltage reflection coefficient of the reflector dish is denoted by $\Gamma_{d}$. Upon first incidence of the sky signal onto the dish, it is reflected from the entire dish and is focussed onto the feed. Hence, for first incidence $\Gamma_{d}=1$. Sky signal that would return to the dish after successive reflections off the feed and the dish is only reflected from a small area around the vertex of the dish and therefore, $\Gamma_{d}<1$. In our analysis, we consider one reflection of the sky power from the feed and subsequent reflection off the dish as one single reflection. If the focal length of the dish is $l$, assuming the feed is located at the focus, the roundtrip delay after first reflection will be $\Delta \tau = {2l \over c}$.

\subsection{Effect of first order reflections on visibility and delay spectrum}
Upon first incidence of the sky signal $\volt_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
\volt_{1} & = & a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp  (1-\Gamma_{f})^{2}\nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber\\
\label{eq4}
\end{eqnarray}
\begin{eqnarray}
\volt_{2} & = & \left [ a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp (1-\Gamma_{f})^{2}\right ] \fngexp \nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} \dfngexp  \right] \fngexp \nonumber\\
\label{eq5}
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna field pattern $a$ and denote them as $a_{f}= a(1-\Gamma_{f})$ here after and define the antenna cross power pattern as $\bm_{f}(\thhat,\nu)=\bmvolt_{f}(\thhat,\nu)\bmvolt_{f}^{*}(\thhat,\nu)$ . Hence, the complex visibility could be,

\begin{eqnarray}
\vis(\vec b, \nu) & = & \int \volt_{1}\volt_{2}^{*} d\Omega \nonumber\\
		        & = & \int \bm_{f} \volt_{sky}|^{2} |\left [ 1+(1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} \dfngexp  \right]|^{2} \fngexp  d\Omega
\label{eq6}		        
\end{eqnarray}

 Furthermore, combining all the terms that contain instrument response we define a common instrument parameter, written as, 
\begin{eqnarray}
\Sigma(\vec b, \nu) & = & \bm_{f} |\left [ 1+ (1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]|^{2}\nonumber\\
 	    & = &  \bm_{f}  \left[ 1+ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} + 2 \Re \left ( (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp \right) \right] 
\label{eq{7}}	    
\end{eqnarray}

@@

Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
 Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay will be located at the same delay $\tau $ in the Fourier domain whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\Delta \tau$. This additional delay extends the upper limit on the maximum geometric delay $\tau_{g}$ to $(\tau_{g}+\Delta \tau)$.
 Thus, the correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal thus extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected. If the dish vertex is so constructed such that the reflection from the dish vertex is negligible i.e, $\Gamma_{d}<<1\ or = 0$ then the second and the third term of the $\Sigma$ vanish. Also, if we assume a $\Gamma_{a} =-10dB$ and $\Gamma_{d}= -20dB$, the second term in the expression of $\Sigma(\vec b, \nu)$ becomes negligible compared to the first and the third term of the expression and, therefore, can be neglected. \\

\subsection{Effect of higher order reflection on the delay spectrum} 
Correlation between the higher order reflections of the signal in any antenna with the direct component of the other would result in a delay response at integral multiple of $\Delta \tau$ and vice versa. 
If $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{equation}
\volt_{1} =  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}]
\end{equation}
and,

\begin{eqnarray}
\volt_{2} & = &  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&& ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \ast \fngexp 
\end{eqnarray}

where $\Phi = 2\pi\nu\Delta \tau$. The relevant terms in the corresponding visibility and subsequently in the delay spectrum would be the ones which represent the correlation between the direct component of one antenna with the reflected components of the other resulting in higher order terms  in $\Sigma(\vec b, \nu)$. The corresponding instrument response may be written as,

\begin{equation}
\Sigma(\vec b, \nu) = a_{f} \left [ 1+ 2 \Re \left [(1-\Gamma_{f}) \Gamma_{f}\Gamma_{d} e^{i\phi}+ (1-\Gamma_{f})^{2}(\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+(1-\Gamma_{f})^{n} (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi} \right ]\right ]
\end{equation}

and the measured visibility spectrum, integrated over all sky would be, 
\begin{equation}\label{eqn:series1}
\vis(\vec b, \nu)  = \int I_{sky}\Sigma(\vec b, \nu) d\Omega
\end{equation}
